---
title: "MSC Basketball Data Analysis"
author: "Ben Peterson"
date: "11/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DataComputing)
library(tidyverse)
library(tigerstats)
```

First, I will load in the full MSC Game Data:

```{r}
load("data/MSCData.rda")
```

## Creating a Column for the Two and Three Point Shooting Percentages

It may be of interest in this project to know what the shooting percentages were for each team in each game. For this reason, I will create the additional columns now.

### Three Point Shooting Percentages

First, I will make the column for the 3 point shooting percentages for each team in each game.

```{r}
MSCDataFinal=data.frame(MSCDataFinal,X3Perc=100*(MSCDataFinal$X3PM/MSCDataFinal$X3PA))
MSCDataFinal
```

### Two Point Shooting Percentages

Next, I will make the column for the 2 point shooting percentages for each team in each game.

```{r}
MSCDataFinal=data.frame(MSCDataFinal,X2Perc=100*(MSCDataFinal$X2PM/MSCDataFinal$X2PA))
MSCDataFinal
```

## MSC Data Analysis

I am creating a data frame called fgs, which will sum the total two and three pointers made and attempted for each team throughout the course of all three seasons. 

```{r}
fgs=data.frame(aggregate(MSCDataFinal$X3PM~MSCDataFinal$team,
                         data=MSCDataFinal,FUN=sum),
               aggregate(MSCDataFinal$X3PA~MSCDataFinal$team,
                         data=MSCDataFinal,FUN=sum),
               aggregate(MSCDataFinal$X2PM~MSCDataFinal$team,
                         data=MSCDataFinal,FUN=sum),
               aggregate(MSCDataFinal$X2PA~MSCDataFinal$team,
                         data=MSCDataFinal,FUN=sum))
fgs=fgs[,c(-3,-5,-7)]

fgs
```

Thus, I can calculate a percentage of shots made for each team in the data frame, and get that team's two and three point average over the three season span.

```{r}
fgs=data.frame(fgs, Perc2=100*(fgs$MSCDataFinal.X2PM/fgs$MSCDataFinal.X2PA))
fgs=data.frame(fgs, Perc3=100*(fgs$MSCDataFinal.X3PM/fgs$MSCDataFinal.X3PA))
fgs
```

## Preliminary Conclusion

So far, it is possible to conclude which teams are the better teams in terms of two point shooting and three point shooting. I will create a table that arranges these in descending order, so that the teams will be ranked in terms of offensive efficiency.

### Two Point Offensive Efficiency Ranking
```{r}
fgs %>%
  select(MSCDataFinal.team, Perc2) %>%
  mutate(ranking=rank(desc(Perc2))) %>%
  arrange(ranking)
```

### Three Point Offensive Efficiency Ranking

```{r}
fgs %>%
  select(MSCDataFinal.team, Perc3) %>%
  mutate(ranking=rank(desc(Perc3))) %>%
  arrange(ranking)
```



### Checking to see how offensive efficiency in two point shots has changed over time

Now, we will analyze the offensive effienciencies for two point shooting for the top 5 ranked teams in the Mid-South Conference, to see how they have changed over time. This will enable readers to see if the teams have improved or gotten worse over the three season span.

```{r}
MSCDataFinal
```


## Graphing the Results

```{r}

```

