---
title: "MSC Basketball Data Analysis"
author: "Ben Peterson"
date: "11/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DataComputing)
library(tidyverse)
library(tigerstats)
```

## Draft Comments

Since you are just getting started I mostly have to address overall structure, leaving many technical details to you as work on the final version of the report.

* Re-read the project assignment pdf:  you must follow all requirements for the draft (format, sections, tables, etc.).  Since data import is a big part of your project, you'll must incorporate at least a discussion of it into your final report.
* As much as possible, do your data wrangling with **dplyr** and other tools we learned in class this semester.  Don't just copy what you did in Spring 2018.
* Do not limit yourself to the analysis you did in Spring 2018.  Here are some suggestions for further work.
    * Now that you have multitple seaosns, you can compute things like FG percentage for each team, for each season.
    * You can choose a few leading teams to follow over all seasons.  For example, you could study a particular offensive stat (FG percentage, say) for each of the leading teams over the years.  You could make a graph using `geom_line()` that shows how each team's FG percentage has varied over the seasons.
* Whatever analysis you do, make sure that it adds up to a coherent investigation, not just a few random tables and graphs.  Build your report around that investigation.  (For example, the Introduction should indicate what questions you bring to the data, and why.)
    
I have a few technical comments below, and also in the R Markdown documents that concerend data cleaning.

## End of Draft Comments

First, I will load in the full MSC Game Data:

```{r}
load("data/MSCData.rda")
```

## Creating a Column for the Two and Three Point Shooting Percentages

It may be of interest in this project to know what the shooting percentages were for each team in each game. For this reason, I will create the additional columns now.

### Three Point Shooting Percentages

First, I will make the column for the 3 point shooting percentages for each team in each game.

**HSW:  In any doucment that is intended to be knit up, do not print out a long data table directl.  Use something like `DT::datatable()` instead.  I'll make those changes here, so I can navigate the knitted document.**

```{r}
MSCDataFinal=data.frame(MSCDataFinal,
                        X3Perc=100*(MSCDataFinal$X3PM/MSCDataFinal$X3PA))
## MSCDataFinal
```

### Two Point Shooting Percentages

Next, I will make the column for the 2 point shooting percentages for each team in each game.

```{r}
MSCDataFinal=data.frame(MSCDataFinal,
                        X2Perc=100*(MSCDataFinal$X2PM/MSCDataFinal$X2PA))
DT::datatable(MSCDataFinal, options = list(
  scrollX = TRUE
))
```

## MSC Data Analysis

I am creating a data frame called fgs, which will sum the total two and three pointers made and attempted for each team throughout the course of all three seasons. 

```{r}
fgs=data.frame(aggregate(MSCDataFinal$X3PM~MSCDataFinal$team,
                         data=MSCDataFinal,FUN=sum),
               aggregate(MSCDataFinal$X3PA~MSCDataFinal$team,
                         data=MSCDataFinal,FUN=sum),
               aggregate(MSCDataFinal$X2PM~MSCDataFinal$team,
                         data=MSCDataFinal,FUN=sum),
               aggregate(MSCDataFinal$X2PA~MSCDataFinal$team,
                         data=MSCDataFinal,FUN=sum))
fgs=fgs[,c(-3,-5,-7)]

DT::datatable(fgs)
```

**HSW:  Ben, you are copying from your Spring 2017 project.  Find an easier way to accomplish the summaries, using dply commands that you learned this semester.**

Thus, I can calculate a percentage of shots made for each team in the data frame, and get that team's two and three point average over the three season span.

```{r}
fgs=data.frame(fgs, Perc2=100*(fgs$MSCDataFinal.X2PM/fgs$MSCDataFinal.X2PA))
fgs=data.frame(fgs, Perc3=100*(fgs$MSCDataFinal.X3PM/fgs$MSCDataFinal.X3PA))
knitr::kable(fgs)
```

## Preliminary Conclusion

So far, it is possible to conclude which teams are the better teams in terms of two point shooting and three point shooting. I will create a table that arranges these in descending order, so that the teams will be ranked in terms of offensive efficiency.

### Two Point Offensive Efficiency Ranking
```{r}
fgs %>%
  select(MSCDataFinal.team, Perc2) %>%
  mutate(ranking=rank(desc(Perc2))) %>%
  arrange(ranking) %>% 
  knitr::kable()
```

**HSW:  Note that when the doucment knits a warning appears here, even though no such warning appears when you run the command from the Console.  It's a known quirk with knitr.  You can suppress such spurious output with  chunk options like `warning = FALSE` and `message = FALSE`.**

### Three Point Offensive Efficiency Ranking

```{r}
fgs %>%
  select(MSCDataFinal.team, Perc3) %>%
  mutate(ranking=rank(desc(Perc3))) %>%
  arrange(ranking) %>% 
  knitr::kable()
```

