---
title: "MSC Game Data"
author: "Ben Peterson"
date: "11/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
<<<<<<< HEAD
library(DataComputing)
=======
library(tidyverse)  # need this (or at least DataComputing) for dplyr stuff
>>>>>>> 1cbbb05a74d824d132a87163fdee90c7efee9f04
```

## Load the Separate Data Frames

Load the data for each year:

```{r}
load("data/S2014.rda")
load("data/S2015.rda")
load("data/S2016.rda")
```


Now, we will combine the season data frames into one data frame containing all
of the games from each season collected.

<<<<<<< HEAD
## Combining the Season Data Frames
=======
## Labeling the Data

First, I will re-label the ID numbers so that they come in order but are separated by season.

```{r eval = FALSE}
S2015$id <- 145:288
S2016$id <- 289:400
```

Do not do the above!  It destroys the point of providing an id number that links each pair of rows associated with a game.



## Combining the Season Data Frames (and adding IDs)
>>>>>>> 1cbbb05a74d824d132a87163fdee90c7efee9f04

First, we will join the 2014, 2015, 2016 data frames into a frame containing all three seasons.

<<<<<<< HEAD
```{r}
MSCData <- rbind(S2014, S2015, S2016)

MSCData <-
+  MSCData %>% 
+  mutate(id = paste(id, Season, sep = "_"))
=======
```{r eval = FALSE}
MSCData <- S2014
for (i in 1:nrow(S2015)){
  MSCData <- S2015[i,]
}
for (i in nrow(S2016)){
  MSCData <- S2016[i,]
}
>>>>>>> 1cbbb05a74d824d132a87163fdee90c7efee9f04
```

Surely it's easier to do this:

```{r}
MSCData <- rbind(S2014, S2015, S2016)
```

Now let's handle those IDs:

```{r}
MSCData <-
  MSCData %>% 
  mutate(id = paste(id, Season, sep = "_"))
```



## Clean the Data

Now do your cleaning (splitting columns into separate variable as needed).

## Save the Data

Then save the full data frame:

```{r}
save(MSCData, file = "data/MSCData.rda")
```





