---
title: "MSC Game Data"
author: "Ben Peterson"
date: "11/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)  # need this (or at least DataComputing) for dplyr stuff
```

## Load the Separate Data Frames

Load the data for each year:

```{r}
load("data/S2014.rda")
load("data/S2015.rda")
load("data/S2016.rda")
```


Now, we will combine the season data frames into one data frame containing all
of the games from each season collected.

## Labeling the Data

First, I will re-label the ID numbers so that they come in order but are separated by season.

```{r eval = FALSE}
S2015$id <- 145:288
S2016$id <- 289:400
```

Do not do the above!  It destroys the point of providing an id number that links each pair of rows associated with a game.



## Combining the Season Data Frames (and adding IDs)

Now that there is no conflict with id labeling, we will join the 2014 and 2015 data
tables, then add the 2016 data table to that.

```{r eval = FALSE}
MSCData <- S2014
for (i in 1:nrow(S2015)){
  MSCData <- S2015[i,]
}
for (i in nrow(S2016)){
  MSCData <- S2016[i,]
}
```

Surely it's easier to do this:

```{r}
MSCData <- rbind(S2014, S2015, S2016)
```

Now let's handle those IDs:

```{r}
MSCData <-
  MSCData %>% 
  mutate(id = paste(id, Season, sep = "_"))
```



## Clean the Data

Now do your cleaning (splitting columns into separate variable as needed).

## Save the Data

Then save the full data frame:

```{r}
save(MSCData, file = "data/MSCData.rda")
```





