---
title: "MSC Game Data"
author: "Ben Peterson"
date: "11/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DataComputing)
library(tidyverse)  # need this (or at least DataComputing) for dplyr stuff
```

## Load the Separate Data Frames

Load the data for each year:

```{r}
load("data/S2014.rda")
load("data/S2015.rda")
load("data/S2016.rda")
```


Now, we will combine the season data frames into one data frame containing all
of the games from each season collected.

## Combining the Season Data Frames (and adding IDs)

First, we will join the 2014, 2015, 2016 data frames into a frame containing all three seasons, separating the game IDs by season.

```{r}
MSCData <- rbind(S2014, S2015, S2016)

MSCData <-
  MSCData %>% 
  mutate(id = paste(id, Season, sep = "_"))
```

##Tidying the MSC Basketball Data

### Field Goal Column Split

Now, I will split the fgm with hyphens into two seperate columns based on whether it was a made shot or an attempted shot -- FGM and FGA.

```{r}
MSCDataFGMSplit<-
  MSCData %>%
  extractMatches("(\\d{2})-(\\d{2})",fgm) %>%
  mutate(FGM=as.numeric(as.character(match1))) %>% 
  mutate(FGA=as.numeric(as.character(match2))) 


MSCDataFGMSplit <- subset(MSCDataFGMSplit, select = -c(fgm, match1, match2))
MSCDataFGMSplit
```

### Free Throw Column Split

Next, I will split the ftm column into two separate columns based on free throws made and free throws attempted -- FTM and FTA

```{r}
MSCDataFTMSplit<-
  MSCDataFGMSplit %>%
  extractMatches(c("(\\d{2})-(\\d{2})","(\\d{1})-(\\d{2})"),ftm) %>%
  mutate(FTM=as.numeric(as.character(match1))) %>%
  mutate(FTA=as.numeric(as.character(match2)))


MSCDataFTMSplit <- subset(MSCDataFTMSplit, select = -c(ftm, match1, match2))
MSCDataFTMSplit
```

### Three Pointer Column Split

Also, I will split the threepm column into two columns, one for three pointers made, and one for three pointers attempted -- X3PM and X3PA

```{r}
MSCDataX3PMSplit<-
  MSCDataFTMSplit %>%
  extractMatches(c("(\\d{1})-(\\d{2})","(\\d{2})-(\\d{2})"),threepm) %>%
  mutate(X3PM=as.numeric(as.character(match1))) %>% 
  mutate(X3PA=as.numeric(as.character(match2))) 


MSCDataX3PMSplit <- subset(MSCDataX3PMSplit, select = -c(threepm, match1, match2))
MSCDataX3PMSplit
```

###

```{r}

MSCDataFinal=data.frame(MSCDataX3PMSplit,X2PM=MSCDataX3PMSplit$FGM-MSCDataX3PMSplit$X3PM,
                   X2PA=MSCDataX3PMSplit$FGA-MSCDataX3PMSplit$X3PA)
MSCDataFinal
```


## Save the Data

Then save the full data frame:

```{r}
save(MSCDataFinal, file = "data/MSCData.rda")
```





